---
title: "Lab 6"
author: "Julianna Renzi"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(sf)
library(tmap)
library(gstat) # krigging
library(stars) # use to make a grid
```

# Spatial krigging

```{r}
# read in data 
# CA depth to groundwater
ca_dgw <- read_sf(here("ca_dgw"), layer = "F2013_DBGS_Points_20150720_093252") %>% # saying to read in the whole layer (anything that starts with that string/can have different extensions)
  clean_names() # lower case snake case
# measurements are in feet
# negative depth is above the ground

# check projection
st_crs(ca_dgw)

# now read in CA counties
# want to read in entire layer so it's all bundled together
ca_counties <- read_sf(here("ca_counties"), layer = "CA_Counties_TIGER2016") %>% 
  clean_names() %>% 
  select(name) # geometry is sticky!

st_crs(ca_counties) # see it's different but has a crs (so don't do set crs, and instead use transform)

ca_counties <- st_transform(ca_counties, st_crs(ca_dgw)) # st is spatio-temporal
# this sets it to the crs of ca_dgw

```

Now plot

```{r}
ggplot() +
  geom_sf(data = ca_counties) +
  geom_sf(data = ca_dgw, aes(color = dgbs)) # dots are observations
```

Make it interactive!

```{r}
tmap_mode("view")

tm_shape(ca_dgw) + # this uses leaflet behind the scenes
  tm_dots("dgbs") # want depths
```

```{r}
# just get one county
sj_county <- ca_counties %>% 
  filter(name == "San Joaquin") # see in geometry have the nodes for the county (single spatial polygon)

# want the values in that county
sj_depth <- ca_dgw %>% 
  st_intersection(sj_county)

plot(sj_depth)
plot(sj_county)

ggplot() +
  geom_sf(data = sj_county) +
  geom_sf(data = sj_depth, aes(color = dgbs))


```

Note pattern BEFORE you do analyses--looks like in the northeast the depth to groundwater increases and also slightly to the southwest

NEED to be worried about "spatial singularities" (i.e. two points at the SAME POINT with different values--cannot compute). Need to remove or alter

```{r}
well_duplicates <- sj_depth %>% 
  get_dupes(latitude, longitude) # find duplicates (from the janitor package)--only is a duplicate if BOTH lat and long are the same
# see there are two points with duplicate information (in reality maybe you'd take the average but here we'll just remove them)

sj_depth <- sj_depth %>% 
  filter(!local_well %in% well_duplicates$local_well)

sj_depth %>% 
  get_dupes(latitude, longitude) # see it worked! No more singularity issue
  
```

## Create variogram

Will help determine weight each point gets for interpolation. Then we need to make a grid that tells R this is where we want predictions made

```{r}
sj_dgw_vgm <- variogram(dgbs ~ 1, data = sj_depth)

plot(sj_dgw_vgm) # see it looks like a variogram (variance increases over distance)
# BUT we want a function not a set of points!
```

```{r}
sj_dgw_vgm_fit <- fit.variogram(sj_dgw_vgm, 
                                model = (vgm(nugget = 20, psill = 3000, range = 30, model = "Gau")
                                  # need estimate of nugget, range, and sill (based on plot above) + tell it which model you want to use (e.g. spherical etc--compare using errors and pick the best, which is Gaussian in this case)
                                ))

sj_dgw_vgm_fit # estimate for nugget, psill (2800), and range (~17)

plot(sj_dgw_vgm, sj_dgw_vgm_fit) # overall looks pretty good, but maybe not as good with high values (but good enough)
# see point of inflection around 17
# bad models might not even converge
```
## Spatial kriging (interpolation)

First need to make a grid for where we want to make predictions

```{r}
# want it only in San Juaquin 
sj_grid <- st_bbox(sj_county) %>% 
  st_as_stars(dx = 0.01, dy = 0.01) %>%  # stars is raster information that exists in multiple layers/bands. dx/dy give resolution of pixels
  st_set_crs(4326) %>% # set reference system
  st_crop(sj_county) # crop to sj_county

plot(sj_grid)
  
```

```{r}
sj_dgw_krige <- krige(dgbs ~ 1, sj_depth, sj_grid, model = sj_dgw_vgm_fit) # from gstat package (function of 1 if we do ordinary kriging--constant but unknown field)

plot(sj_dgw_krige) # stored as var1.pred (predicted)--also stores variance for each prediction
```

